Gamma-Ray Burst Afterglows In The Multi-Messenger Era: Numerical Models and Closure Relations

### 2019-09-25 structuredJets.tex
Arxiv Version??

%\newpage
%
%\subsection{Old Draft}
%
%The binary neutron star merger \gwbns{} and its associated electromagnetic counterparts provided a wealth of data to high energy astrophysics (CITE).  Not least among these was its peculiar non-thermal afterglow, first detected in X-rays 9 days after the merger \citep{Troja:2017aa}.  This was followed by radio 16 days after the merger and optical 73 days after the merger (CITE Radio \& Optical follow-up).  Unlike any short gamma-ray burst (GRB) afterglow observed before, \grbbns{} steadily {\em rose} in brightness for ~160 days in all observed bands.  After peak, the emission quickly decayed at a rate similar to other GRB afterglows.
%
%The absence of early X-ray emission, with early non-detections by \swiftXRT{} and a deep upper limit from \chandra{}, led early analyses to consider the effect if viewing angle on the emission \citep{Troja:2017aa} (CITE MORE).  Given the gravitational wave signal of a binary neutron star merger is observable almost regardless of of orientation, it is in fact expected that GRB/GW events would be viewed at moderate inclination.
%
%As \grbbns{}'s afterglow continued to rise for weeks after the burst, it became evident the emission could not be explained by a simple uniform ``top hat'' jet viewed off axis.  More sophisticated models were considered: ``structured jets'' with non-uniform angular distribution of energy and quasi-isotropic ``refreshed shocks'' with significant velocity stratification of the ejecta and modest Lorentz factors.  The observation of apparent superluminal motion of the radio centroid and the steep turnover of the light curve after 160 days confirmed the relativistic and collimated nature of the outflow, in accordance with the structured jet picture (CITE Mooley, Ghirlanda, YITL, other steepening paper).
%
%It is expected, of course, that 
%
%Structured jets, blast waves with a non-uniform angular distribution of energy, have been invoked to explain several temporal features in GRB afterglow light curves.  
%
%\subsubsection{History of structure and viewing angle}
%
%The first GRB afterglow was discovered by \beppoSAX{} as part of GRB 970228 \citep{Wijers:1997aa}.  GRB afterglows present as broadband synchrotron emission following the prompt gamma-ray signal and can last for days, weeks, months, and years.  
%
%
%\subsubsection{What we're doing}
%
%In this paper we develop a theoretical model of the electromagnetic afterglow of generic structured jets for arbitrary viewing inclinations.  The model takes as input the energy and Lorentz factor angular profiles, which may be intrinsic to the jet launching mechanism 
%
%The light curves resemble broken power laws.  We calculate viewing angle dependent closure relations, directly relating the effective temporal evolution of the light curve to the viewing angle and explosion geometry.  
%
%
%\subsection{Old Old Draft}
%
%The electromagnetic afterglow of \gwbns{} gave, quite literally, a new view on gamma-ray burst (GRB) afterglows.  
%
%A canonical afterglow was almost immediately ruled out \citep{Troja:2017aa}.  
%
%
%The structure of the blast wave giving rise to gamma-ray burst (GRB) afterglows has been a subject of interest since their discovery by \beppoSAX{} in GRB 970228 \citep{Wijers:1997aa}.
%
%{\color{blue} LUIGI: ?was found since early afterglow observations with BeppoSAX, when the
%presence of a structured outflow has been considered for explaining some of the
%temporal breaks observed in the light curves of a number of GRBs (Panaitescu and
%Kumar 2003, Apj 592: 390, Panaitescu 2005 MNRAS 362, 921). Candidates for
%structured jet exhibit a variation of the decay index at the break smaller that that
%expected for a top-hat spreading jet ($\Delta \alpha \approx 0.8$) and larger than
%that induced by the crossing of the cooling frequency ($\Delta\alpha=0.25$). Energy
%injection by a long lasting central engine can also induce temporal breaks, either
%flattening or steepening. Other effects such as transition to non-relativistic
%expansion, the transition from wind-like to ISM (Piro et al. 2005, Apj 623 314) or the
%emergence of an Inverse compton component (Corsi \& Piro
%2006A\&A...458..741C) can also produce breaks, in these case associated with a
%flattening of the light curve.
%More recenlty Ryan et al (2015) presented evidence of a structured jet in a sample
%based on SWIFT observations. Considering the number of processes that can
%produce breaks mentioned above, no undisputable observational evidence of a
%structured jet became available till the observations of GW170817. Real jets}
%
%
%Gamma-ray burst (GRB) afterglows, since their discovery by \beppoSAX{} in GRB 970228, have been a potent tool for studying GRBs and their progenitors.
%
%
%
%Gamma-ray burst (GRB) afterglows are commonly modeled as top hat blast waves with no internal structure pointed directly at the observer. This has been largely sufficient given the available data, although some evidence of viewing angle effects has been found \citep{Ryan:2015aa}.  New observational capabilities have witnessed new phases of afterglow evolution, such as the discovery of plateaus with Swift.  Real jets (of course) have internal structure, but the top hat assumption has been a sufficiently spherical cow. 
%
%\gwbns{} changes the picture: new object with a strange afterglow.  Steep/late rise, LONG slow rise, turnover at 160d.  The slow rise is impossible to reproduce with standard afterglow evolution: first observation of a \emph{structured} phase of evolution. 
%
%The dominant contribution may be an angular distribution of energy within the blast wave (``structured jet'') combined with off-axis viewing (CITE Rossi04).  For \gwbns{} this corresponds to the ``successful jet'' picture.  Alternately, radial stratification of velocity (``refreshed shock'') can cause a similar phase, most simply with a quasi-spherical (or at least top-hat) angular structure.  This situation is produced by the ``choked jet''  or ``cocoon'' picture of \gwbns{}. Understanding the effects of both structures on afterglows, and their dependence on the parameters of the burst, are vital to understanding GW counterparts.
%
%To this end, we have calculated the power law temporal slope of the structured phase in both angular and radial structure dominated cases, along with the corresponding observed break times.  We have developed a semi-analytic numerical model of structured GRB afterglows for detailed comparison to data, which is now publicly available.
%
%Summary: In section 2 we discuss structured jets, in section 3 refreshed shocks, in section 4 our numerical model, section 5 application to \gwbns{}, and section 6 future applications and discussion.

@@

The \emph{structured jet} is a GRB jet model where the isotropic-equivalent energy of the blast wave is a function of the angle from the jet axis: $\Eiso = E(\theta)$. The particular angular structure of a jet is first imposed by the jet launching mechanism and then modified by the sculpting that occurs as the jet burrows out of the surrounding ejecta debris (as in a binary neutron star merger) or stellar envelope (as in a collapsar).  In the case of neutron star mergers, it is not a given that there is sufficient ejecta in the polar regions to significantly alter the intrinsic jet structure.

@@

% TODO: Look at this again, more detail.
The rest-frame synchrotron emissivity $\epsilon'_{\nu'}$ may be calculated with varying levels of sophistication.  We use the standard broken power law formalism with characteristic frequencies $\nu_m$ and $\nu_c$, with the cooling frequency $\nu_c$ calculated via the global cooling approximation \citep{Granot:2002aa, van-Eerten:2010aa}.  The fluid behind the shock has rest-frame number density $n' = 4n_0\gamma$, thermal energy $e'_{th} = (\gamma-1)m_p n' c^2$, and magnetic field strength $B = \sqrt{8\pi e'_{th} \epsB}$, where $\epsB$ is the fraction of thermal energy in the magnetic field. We assume a fraction $\xiN$ of the electron population is shock-accelerated electrons with a fraction $\epse$ of the thermal energy and have a Lorentz factor distribution $N(\gamma_e) \propto \gamma_e^{-p}$ with index $p>2$.

@@

%These errors quoted are numerical errors associated with the schemes used to solve Equations \ref{eq:Rdot}, \ref{eq:E}, and \ref{eq:thetadot} and perform the integral in Equation \ref{eq:flux2}.  That is, they are the errors relative to solving these equations perfectly.  The equations themselves are, however, an approximation themselves to the full physics of an afterglow light curve.  Comparison and calibration against analytic calculations and hydrodynamic simulations is necessary to ensure physically realistic results.  
%
\begin{figure*}
	\includegraphics[width=\textwidth]{figs/boxfit_comp.pdf}
	\caption{Top Panel: Comparison between top hat jet light curves from \afterglowpy{} (solid lines) and \boxfit{} (dashed lines). Bottom Panel: Fractional difference between \afterglowpy{} and \boxfit{} light curves.  Four representative light curves are shown: radio on-axis ($\thobs=0$, $\nu=10^9$Hz, blue), radio off-axis ($\thobs=0.16$rad, $\nu=10^9$Hz, orange), X-ray on-axis ($\thobs=0$, $\nu=10^{18}$Hz, green), and X-ray off-axis ($\thobs=0.16$rad, $\nu=10^{18}$Hz, red).  Remaining parameters are shared: $\thC=0.1$rad, $\Eiso=10^{52}$erg, $n_0 = 10^{-3}$cm$^{-3}$, $p=2.2$, $\epse=10^{-1}$, $\epsB=10^{-2}$, $\dL=3.09 \times 10^{26}$cm, $z=0.028$. \label{fig:boxfitComp}}
%TODO: put distance in Mpc
\end{figure*}

Figure \ref{fig:boxfitComp} shows a comparison between top-hat light curves calculated with \afterglowpy{} and \boxfit{}.   Figure \ref{fig:boxfitComp} shows the light curve of a $\thC=0.1$ rad top-hat jet at radio and X-ray frequencies ($10^9$ Hz and $10^{18}$ Hz, respectively) with aligned ($\thobs = 0$) and misaligned ($\thobs=0.16$ rad) viewing angles.  The light curves begin in the ultra relativistic Blandford-Mckee phase and continue to the Newtonian Sedov phase.

The overall agreement is good: \afterglowpy{} captures the salient features of both aligned and misaligned light curves including the jet break, spectral shape, transition to the Sedov phase, and the early steep rise for misaligned viewing.  
At early times, \boxfit{} lacks the simulation coverage to produce accurate fluxes, leading to a large discrepancy.  Once proper coverage is attained, \boxfit{} and \afterglowpy{} show less than 50\% relative discrepancy until the jet break passes and jet spreading begins.  After the jet break the \afterglowpy{} light curves show a marked excess relative to \boxfit{}, the relative error increasing to approximately 1.0 and 2.0 for the aligned and misaligned radio light curves, respectively.  The X-ray light curves achieve approximately half the error of the radio.  After a short period the \afterglowpy{} light curves begin to decrease, recovering the \boxfit{} slope, and asymptote to similar light curves in the Newtonian regime with a small constant offset from \boxfit{}.

As expected, \afterglowpy{} models the early and late afterglow well, but is somewhat less accurate in an intermediate phase during jet spreading.  In particular it appears that \afterglowpy{} begins its spreading late relative to \boxfit{}.  This phase is precisely where the non-linear hydrodynamics of the blast wave are most important, and hardest to model with simple semi-analytic approximations.  Any work utilizing \afterglowpy{} should be aware of this fact, and be careful when treating data from this regime.

@@

% TODO Finish this paragraph
To fully characterize the light curves of structured jets we use \afterglowpy{} to construct light curves exploring dependence on jet structure model, viewing angle $\thobs$, opening angle $\thC$, and synchrotron regime.

@@

\begin{deluxetable*}{CCCCCCCCC}
	\tablecaption{Structured jet closure relations: $F_\nu \propto \tobs^{\alpha} \nuobs^{\beta}$. The ``far-off-axis'' phase ($\alpha_{\mathrm{far-off-axis}}$)occurs at early times and only if there is no material directly in the observer's line of sight ($\thobs > \thW$). The ``generic'' slope ($\alpha_{\mathrm{generic}}$) applies whenever the jet is relativistic, non-spreading, and the observer is within the beaming cone of some part of the jet ($|\thobs - \theta_*| < \gamma^{-1}(\theta_*)$). The $g$ parameter is positive and depends on the viewing angle $\thobs$ and the specific structure profile $E(\theta)$.  See Equations \eqref{eq:gdef} and \eqref{eq:geff} for the definition and effective values of $g$.  For a top hat jet $g=0$ and for a Gaussian jet $g \approx \thobs^2/(4\thC^2)$. The $\som$ parameter encodes how the effective angular size of the jet is changing with Lorentz factor: $\Delta \Omega \propto \gamma^{-\som}$.  The ``structured'' phase ($\alpha_{\mathrm{struct}}$) is seen by misaligned viewers ($\thobs > \thC$), occurs before the jet break, and corresponds to $\som \approx 1$. The ``pre-jet break'' phase ($\alpha_{\mathrm{pre}}$) is seen by aligned viewers ($\thobs \lesssim \thC$) and by early misaligned viewers if there is high Lorentz factor material in their line of sight.  The first ``post-jet break'' slope ($\alpha_{\mathrm{post, 1}}$) given here does not include the effects of jet spreading and as such is shallower than the true post-jet break slope.  The second ``post-jet break'' slope ($\alpha_{\mathrm{post, 2}}$) was calculated in \citet{Sari:1999aa}, corresponds to a jet undergoing exponential spreading, and is steeper than the true post-jet break slope.  \label{tab:slopes}}
	\tablehead{\colhead{Regime}& \colhead{Label} & \colhead{$\alpha_{\mathrm{far-off-axis}}$} & \colhead{$\alpha_{\mathrm{generic}}$} & \colhead{$\alpha_{\mathrm{struct}}$}& \colhead{$\alpha_{\mathrm{pre}}$}& \colhead{$\alpha_{\mathrm{post, 1}}$} & \colhead{$\alpha_{\mathrm{post, 2}}$} & \colhead{$\beta$}}
	\startdata
	& & & & \som = 1 & \som=2 & \som=0 &\\
	& & & &  & g=0 & g=0 & \\[0cm]
	\nu<\nu_m<\nu_c     & D	& 7 		& \dfrac{-2 + 3\som + 3g}{8+g} 		& \dfrac{1 + 3g}{8+g} 	&1/2		& -1/4	& -1/3 & 1/3 \\[0.4cm]
	\nu < \nu_c < \nu_m & E	& 17/3 	& \dfrac{-14/3 + 3\som + 11g/3}{8+g}  & \dfrac{-5/3 + 11g/3}{8+g} &1/6 	& -7/12	& -1    & 1/3 \\[0.4cm]
	\nu_c < \nu < \nu_m & F	& 13/2 	& \dfrac{-8 + 3\som + 2g}{8+g} 		& \dfrac{-5 + 2g}{8+g} 	& -1/4 	& -1		& -1	  & -1/2 \\[0.4cm]
	\nu_m<\nu<\nu_c     & G 	&\dfrac{15 - 3p}{2} & \dfrac{-6p + 3\som + 3g}{8+g} 	& \dfrac{3 - 6p +3g}{8+g} 	& -\dfrac{3(p-1)}{4} & -\dfrac{3p}{4}	& -p	  & \dfrac{1-p}{2} \\[0.4cm]
	\nu_m, \nu_c < \nu & H 	& \dfrac{16 - 3p}{2} 	& \dfrac{-6p - 2 + 3\som + 2g}{8+g} 	& \dfrac{1-6p  +2g}{8+g} 	& -\dfrac{3p-2}{4} & -\dfrac{3p+1}{4} & -p & -\dfrac{p}{2} \\[0.4cm]
	\enddata
\end{deluxetable*}

\begin{figure*}
	\includegraphics[width=\textwidth]{figs/lc_phase_sketch.pdf}
	\caption{Illustration of possible synchrotron light curves from the forward shock of a structured jet.  The specific order of light curve phases depends on where the viewer sits, $\thobs$, relative to the jet half opening angle $\thC$ and the outer truncation angle $\thW$. A \emph{coasting} phase, ending at $\tdec$, is assumed to have already completed.  Each panel shows a fiducial \afterglowpy{} light curve (solid blue lines), power law segments $t^\alpha$ with slopes from the closure relations in Table \ref{tab:slopes} (black dashed lines), and representative break times between each phase of evolution (grey dashed vertical lines). The two dashed lines in the ``post-jet break'' section bracket the possible light curve behaviour.  The viewing angle $\thobs = 0$ rad, $0.2$ rad, and $0.4$ rad in the left, center, and right panels respectively.  Remaining parameters are shared: $\thC=0.05$ rad, $\thW=0.3$ rad, $\Eiso=10^{53}$ erg, $n_0 = 10^{-3}$ cm$^{-3}$, $p=2.2$, $\epse=10^{-1}$, $\epsB=10^{-3}$, $\dL=10^{28}$ cm, $z=0.5454$. \label{fig:phaseSketch}}
\end{figure*}

\emph{Coasting} --- Before the blast wave begins to decelerate it coasts at constant Lorentz factor. The emissivity is constant, and the total flux increases $F_\nu \propto t^3$.

@@

An alternative mechanism to produce slow decays or rises in afterglow light curves is the \emph{refreshed shock}, where velocity stratification of the ejecta causes a prolonged period of energy injection in the afterglow \citep{Rees:1998aa, Panaitescu:1998aa, Sari:2000aa}.  This material may be a tail of fast outflowing ejecta or ejecta material accelerated through interaction with a possibly choked jet (e.g. a cocoon) \citep{Nakar:2011aa, Hotokezaka:2015aa}.  In either case, slow material initially coasting behind the shock is gradually incorporated into the blast wave as it decelerates, ``refreshing'' the shock.  This mechanism is a general scenario for energy injection \citep{Zhang:2006aa}, and was proposed as the mechanism behind the \gwbns{} afterglow in the choked-jet scenario (e.g. \citep{Mooley:2018aa, Nakar:2018aa}).

In the simplest case the visible part of the blast wave is assumed to be quasi-spherical.  The velocity distribution of material behind the blast wave is specified by $E(u)$, the energy of all material in the ejecta with four-velocity greater than $u$.  This is typically taken to be a power law in $u$ within the finite domain $[\umin, \umax]$:
\begin{equation}
	E(u) = E_r u^{-k}\ , \label{eq:Eu}
\end{equation}
where $u \in (\umin, \umax)$ is the dimensionless four-velocity, $k>0$ is the power law index, and $E_r$ is a normalization factor.  The mass ejected with velocity $\umax$ is $\Mej$.

%
%  TODO:  Comment: define tdec, tmin IN THE TEXT.   Add (brief) discussion of light curve phases.
%

% TODO: Put in obs frame?

A blast wave will have an initial coasting period before sweeping up enough mass in the ambient medium to begin deceleration and be subject to refreshed shocks.  The transition between coasting and decelerating occurs in the bursters frame at $\tdec$. The shock refreshment ends when the blast wave decelerates past $\umin$, in the observers frame at $\tumin$.  In the single shell model with trans-relativistic equation of state used here, these times are:
\begin{align}
	\tdec &= \left(\frac{9}{4\pi} \frac{\Mej}{n_0 \Mp c^3} \frac{\gmax^2}{\bmax^3(\gmax+1)(4\umax^2+3)} \right)^{1/3}  \ ,\label{eq:tdec} \\
	\tumin &= \left(\frac{9}{16\pi} \frac{E_r}{n_0 \Mp c^5}\right)^{1/3} \umin^{-8/3} \ .\label{eq:tumin}
\end{align}
Equation \eqref{eq:tumin} assumes $E_r \gg \Mej c^2$.   The temporal and spectral slopes in this mechanism depend on the energy distribution index $k$ and are given in Table \ref{tab:refreshSlopes}.

@@

The mapping between energy injection and refreshed shocks was done by \cite{Zhang:2006aa}.  That work specifies the index $s$ of the mass distribution $M_{>u} \propto u^{-s}$, which is related simply to the index $k$ by $k = s-1$. Under this mapping the closure relations in Table \ref{tab:slopes} are equivalent to those in \cite{Zhang:2006aa} and the model can be seen as equivalent to an energy injection model with $q = (8-2k) / (8+k)$.

Misaligned structured jets do not map as neatly onto energy injection models as the scaling of the visible patch, $\som = 1.0$, is different than in standard afterglows, $\som = 2$. Once this change is accounted for, the $g$ parameter plays the same role as $k$, as can be seen by a simple comparison between $\alpha_{\text{struct}}$ in Table \ref{tab:slopes} and $\alpha_{\text{refresh}}$ in Table \ref{tab:refreshSlopes}.  That is, a structured jet light curve is equivalent to a standard energy injection model with:
\begin{align}
	q &= (8-2g) / (8+g)\ ,  \label{eq:struct2ei} \\
	\som &= 1\ .
\end{align}

@@

This degeneracy is a blessing and a curse.  The presence of an extended slowly decaying or rising afterglow is a largely model independent prediction of misaligned viewing and the presence of non-trivial jet structure.  Unfortunately, this same model independence makes it very difficult to distinguish between jet structure profiles, leaving both $E(\theta)$ and $\thobs$ uncertain.

>>>>

@@

The \emph{structured jet} is a GRB jet model where the isotropic-equivalent energy of the blast wave is a function of the angle from the jet axis: $\Eiso = 4\pi dE/d\Omega \equiv E(\theta)$. The particular angular structure of a jet is first imposed by the jet launching mechanism and then modified by the sculpting that occurs as the jet burrows out of the surrounding ejecta debris (as in a binary neutron star merger) or stellar envelope (as in a collapsar).  In the case of neutron star mergers, it is not a given that there is sufficient ejecta in the polar regions to significantly alter the intrinsic jet structure.

@@

The rest-frame synchrotron emissivity $\epsilon'_{\nu'}$ may be calculated with varying levels of sophistication.  We use the standard broken power law formalism with characteristic frequencies $\nu_m$ and $\nu_c$, with the cooling frequency $\nu_c$ calculated via the global cooling approximation \citep{Granot:2002aa, van-Eerten:2010aa}.  The fluid behind the shock has rest-frame number density $n' = 4n_0\gamma$, thermal energy $e'_{th} = (\gamma-1)m_p n' c^2$, and magnetic field strength $B = \sqrt{8\pi e'_{th} \epsB}$, where $\epsB$ is the fraction of thermal energy in the magnetic field. We assume a fraction $\xiN$ of the electron population is shock-accelerated electrons with a fraction $\epse$ of the thermal energy and have a Lorentz factor distribution $N(\gamma_e) \propto \gamma_e^{-p}$ with index $p>2$.

@@

%These errors quoted are numerical errors associated with the schemes used to solve Equations \ref{eq:Rdot}, \ref{eq:E}, and \ref{eq:thetadot} and perform the integral in Equation \ref{eq:flux2}.  That is, they are the errors relative to solving these equations perfectly.  The equations themselves are, however, an approximation themselves to the full physics of an afterglow light curve.  Comparison and calibration against analytic calculations and hydrodynamic simulations is necessary to ensure physically realistic results.  
%
\begin{figure*}
	\includegraphics[width=\textwidth]{figs/boxfit_comp.pdf}
	\caption{Top Panel: Comparison between top hat jet light curves from \afterglowpy{} (solid lines) and \boxfit{} (dashed lines). Bottom Panel: Fractional difference between \afterglowpy{} and \boxfit{} light curves.  Four representative light curves are shown: radio aligned ($\thobs=0$, $\nu=10^9$Hz, blue), radio misaligned ($\thobs=0.16$rad, $\nu=10^9$Hz, orange), X-ray aligned ($\thobs=0$, $\nu=10^{18}$Hz, green), and X-ray misaligned ($\thobs=0.16$rad, $\nu=10^{18}$Hz, red).  Remaining parameters are shared: $\thC=0.1$rad, $\Eiso=10^{52}$erg, $n_0 = 10^{-3}$cm$^{-3}$, $p=2.2$, $\epse=10^{-1}$, $\epsB=10^{-2}$, $\dL=3.09 \times 10^{26}$cm, $z=0.028$. \label{fig:boxfitComp}}
\end{figure*}

Figure \ref{fig:boxfitComp} shows a comparison between top-hat light curves calculated with \afterglowpy{} and \boxfit{}.   Figure \ref{fig:boxfitComp} shows the light curve of a $\thC=0.1$ rad top-hat jet at radio and X-ray frequencies ($10^9$ Hz and $10^{18}$ Hz, respectively) with aligned ($\thobs = 0$) and misaligned ($\thobs=0.16$ rad) viewing angles.  The light curves begin in the ultra relativistic Blandford-McKee phase and continue to the Newtonian Sedov phase.

The overall agreement is good: \afterglowpy{} captures the salient features of both aligned and misaligned light curves including the jet break, spectral shape, transition to the Sedov phase, and the early steep rise for misaligned viewing.  
At early times, \boxfit{} lacks the simulation coverage to produce accurate fluxes, leading to a large discrepancy.  Once proper coverage is attained \boxfit{} and \afterglowpy{} show less than 50\% relative discrepancy, apart from the on-axis radio light curve which briefly has a 60\% discrepancy at the time of the jet break ($t_b \approx 7 \times 10^5$ s for the aligned light curves, $t_b \approx 3\times 10^6$ s for the misaligned).  After the jet break the \afterglowpy{} light curves recover the \boxfit{} slope, and asymptote to similar light curves in the Newtonian regime with a small constant offset from \boxfit{}.

As expected, \afterglowpy{} models the early and late afterglow well, but is somewhat less accurate in the intermediate phase during the onset of jet spreading. This phase is precisely where the non-linear hydrodynamics of the blast wave are most important, and hardest to model with simple semi-analytic approximations.  Any work utilizing \afterglowpy{} should be aware of this fact, and be careful when treating data from this regime.

@@

Afterglow light curves, even when viewed off axis, can typically be described as a piecewise set of power law segments $F_\nu \propto \tobs^\alpha \nuobs^\beta$, each denoting particular phases of evolution of the blast wave and its emitting region.  Particular phases have been identified empirically \citep{Nousek:2006aa} and theoretically \citep{Zhang:2006aa}.   Within a particular phase, the slopes $\alpha$ and $\beta$ are often related through a \emph{closure relation}: an equality determining $\alpha$ from $\beta$ or vice versa.  Phases change at \emph{break times} when the temporal slope $\alpha$ (and potentially the spectral slope $\beta$) transition rapidly from the previous phase to the next.

To fully characterize the light curves of structured jets we use \afterglowpy{} to construct light curves exploring dependence on jet structure model, viewing angle $\thobs$, opening angle $\thC$, and synchrotron regime.  We identify the relevant phases in the afterglow and present new closure relations and jet break time scalings for the \emph{structured} phase of evolution: when the patch of the blast wave dominating the emission is sliding from the line of sight towards the jet core.

@@

\begin{deluxetable*}{CCCCCCCCC}
	\tablecaption{Structured jet closure relations: $F_\nu \propto \tobs^{\alpha} \nuobs^{\beta}$. The ``far-off-axis'' phase ($\alpha_{\mathrm{far-off-axis}}$) occurs at early times and only if there is no relativistic material directly in the observer's line of sight ($\thobs > \thW$). The ``generic'' slope ($\alpha_{\mathrm{generic}}$) applies whenever the jet is relativistic, non-spreading, and the observer is within the beaming cone of some part of the jet ($|\thobs - \theta_*| < \gamma^{-1}(\theta_*)$). The $g$ parameter is positive and depends on the viewing angle $\thobs$ and the specific structure profile $E(\theta)$.  See Equations \eqref{eq:gdef} and \eqref{eq:geff} for the definition and effective values of $g$.  For a top hat jet $g=0$ and for a Gaussian jet $g \approx \thobs^2/(4\thC^2)$. The $\som$ parameter encodes how the effective angular size of the jet is changing with Lorentz factor: $\Delta \Omega \propto \gamma^{-\som}$.   The ``pre-jet break'' phase ($\alpha_{\mathrm{pre}}$) is seen by aligned viewers ($\thobs \lesssim \thC$) and by early misaligned viewers if there is high Lorentz factor material in their line of sight.  The ``structured'' phase ($\alpha_{\mathrm{struct}}$) is seen by misaligned viewers ($\thobs > \thC$), occurs before the jet break, and corresponds to $\som \approx 1$.   The ``post-jet break'' slopes ($\alpha_{\mathrm{post, 1}}$ and $\alpha_{\mathrm{post, 2}}$) bracket the possible post-jet break behaviour.  The first is for a non-spreading jet while the second corresponds to a jet undergoing exponential spreading and was calculated \citet{Sari:1999aa}.\label{tab:slopes}}
	\tablehead{\colhead{Regime}& \colhead{Label} & \colhead{$\alpha_{\mathrm{far-off-axis}}$} & \colhead{$\alpha_{\mathrm{generic}}$} & \colhead{$\alpha_{\mathrm{pre}}$} & \colhead{$\alpha_{\mathrm{struct}}$}& \colhead{$\alpha_{\mathrm{post, 1}}$} & \colhead{$\alpha_{\mathrm{post, 2}}$} & \colhead{$\beta$}}
	\startdata
	& & & & \som = 2 & \som=1 & \som=0 &\\
	& & & &  g = 0 & & g=0 & \\[0cm]
	\nu<\nu_m<\nu_c     & D	& 7 		& \dfrac{-2 + 3\som + 3g}{8+g} 		&1/2	& \dfrac{1 + 3g}{8+g} 		& -1/4	& -1/3 & 1/3 \\[0.4cm]
	\nu < \nu_c < \nu_m & E	& 17/3 	& \dfrac{-14/3 + 3\som + 11g/3}{8+g}  &1/6 & \dfrac{-5/3 + 11g/3}{8+g} 	& -7/12	& -1    & 1/3 \\[0.4cm]
	\nu_c < \nu < \nu_m & F	& 13/2 	& \dfrac{-8 + 3\som + 2g}{8+g} 		& -1/4 & \dfrac{-5 + 2g}{8+g} 		& -1		& -1	  & -1/2 \\[0.4cm]
	\nu_m<\nu<\nu_c     & G 	&\dfrac{15 - 3p}{2} & \dfrac{-6p + 3\som + 3g}{8+g} & -\dfrac{3(p-1)}{4}	& \dfrac{3 - 6p +3g}{8+g} 	 & -\dfrac{3p}{4}	& -p	  & \dfrac{1-p}{2} \\[0.4cm]
	\nu_m, \nu_c < \nu & H 	& \dfrac{16 - 3p}{2} 	& \dfrac{-6p - 2 + 3\som + 2g}{8+g} 	& -\dfrac{3p-2}{4} & \dfrac{1-6p  +2g}{8+g} 	& -\dfrac{3p+1}{4} & -p & -\dfrac{p}{2} \\[0.4cm]
	\enddata
\end{deluxetable*}

\begin{figure*}
	\includegraphics[width=\textwidth]{figs/lc_phase_sketch.pdf}
	\caption{Illustration of possible synchrotron light curves from the forward shock of a structured jet.  The specific order of light curve phases depends on where the viewer sits, $\thobs$, relative to the jet half opening angle $\thC$ and the outer truncation angle $\thW$. A \emph{coasting} phase, ending at $\tdec$, is assumed to have already completed.  Each panel shows a fiducial light curve (solid blue lines), power law segments $t^\alpha$ with slopes from the closure relations in Table \ref{tab:slopes} (black dashed lines), and representative break times between each phase of evolution (grey dashed vertical lines). The two dashed lines in the ``post-jet break'' section bracket the possible light curve behaviour.  The viewing angle $\thobs = 0$ rad, $0.2$ rad, and $0.4$ rad in the left, center, and right panels respectively.  Remaining parameters are shared: $\thC=0.05$ rad, $\thW=0.3$ rad, $\Eiso=10^{53}$ erg, $n_0 = 10^{-3}$ cm$^{-3}$, $p=2.2$, $\epse=10^{-1}$, $\epsB=10^{-3}$, $\dL=10^{28}$ cm, $z=0.5454$. 
	\label{fig:phaseSketch}}
\end{figure*}

\emph{Coasting} --- Before the blast wave begins to decelerate it coasts at constant Lorentz factor. The flux depends only on the total volume of the emitting region and the evolution of the cooling frequency $\nu_c$.

@@

An alternative mechanism to produce slow decays or rises in afterglow light curves is the \emph{refreshed shock}, where velocity stratification of the ejecta causes a prolonged period of energy injection in the afterglow \citep{Rees:1998aa, Panaitescu:1998aa, Sari:2000aa}.  This material may be a tail of fast outflowing ejecta or ejecta material accelerated through interaction with a possibly choked jet (e.g. a cocoon) \citep{Nakar:2011aa, Hotokezaka:2015aa}.  Both scenarios are particularly relevant to afterglows arising from binary neutron star mergers.  In either case, slow material initially coasting behind the shock is gradually incorporated into the blast wave as it decelerates, ``refreshing'' the shock.  This mechanism is a general scenario for energy injection \citep{Zhang:2006aa}, and was proposed as the mechanism behind the \gwbns{} afterglow in the choked-jet picture (e.g. \citep{Mooley:2018aa, Nakar:2018aa}).

In the simplest case the visible part of the blast wave is assumed to be quasi-spherical.  The velocity distribution of material behind the blast wave is specified by $E(u)$, the energy of all material in the ejecta with four-velocity greater than $u$.  This is typically taken to be a power law in $u$ within the finite domain $[\umin, \umax]$:
\begin{equation}
	E(u) = E_0 \left(\frac{u}{\umax}\right)^{-k}\ , \label{eq:Eu}
\end{equation}
where $u \in (\umin, \umax)$ is the dimensionless four-velocity, $k>0$ is the power law index, and $E_0$ the kinetic energy of the fastest material with $u=\umax$ initially.  The mass ejected with velocity $\umax$ is $\Mej = E_0 / ((\gmax -1)c^2)$.

A blast wave will have an initial coasting period before sweeping up enough mass in the ambient medium to begin deceleration and be subject to refreshed shocks.  The transition between coasting and decelerating occurs in the bursters frame at $\tdec$. The shock refreshment ends at $\tumin$ when the blast wave decelerates past $\umin$, when there is no longer any material to refresh it.   At this point the afterglow transitions to a standard pre-jet break phase. In the single shell model with in the limit $\umax \gg 1$, these transition times (in the observer's frame) are:
\begin{align}
	\tdec &= \frac{1}{4}(1+z)   \umax^{-8/3}\ \! \tNR  \ , \label{eq:tdec} \\ %\left(\frac{9}{4\pi} \frac{\Mej}{n_0 \Mp c^3} \frac{\gmax^2}{\bmax^3(\gmax+1)(4\umax^2+3)} \right)^{1/3}  \ ,\label{eq:tdec} \\
	\tumin &= \frac{1}{4}(1+z) \frac{2+k}{8+k} \left(\frac{\umax}{\umin}\right)^{(8+k)/3} \umax^{-8/3}\ \! \tNR \ .\label{eq:tumin}
\end{align}
 The temporal and spectral slopes for the coasting and refreshing phases are given in Table \ref{tab:refreshSlopes}.  The refreshing phase slope $\alpha_{\mathrm{refresh}}$ depends on the energy index $k$.  A quasi-spherical blast wave experiences very little or no jet spreading and should smoothly transition to the Newtonian Sedov phase after $\tumin$ passes and the energy injection ends.

@@

The mapping between energy injection and refreshed shocks was done by \cite{Zhang:2006aa}.  That work specifies the index $s$ of the mass distribution $M_{>u} \propto u^{-s}$, which is related simply to the index $k$ by $k = s-1$. Under this mapping the closure relations in Table \ref{tab:refreshSlopes} are equivalent to those in \cite{Zhang:2006aa} and the model can be seen as equivalent to an energy injection model with $q = (8-2k) / (8+k)$.

Misaligned structured jets do not map as neatly onto energy injection models as the scaling of the visible patch, $\som = 1.0$, is different than in standard afterglows, $\som = 2$. Once this change is accounted for, the $g$ parameter plays the same role as $k$, as can be seen by a simple comparison between $\alpha_{\text{struct}}$ in Table \ref{tab:slopes} and $\alpha_{\text{refresh}}$ in Table \ref{tab:refreshSlopes}.  That is, a structured jet light curve is similar to a standard energy injection model with:
\begin{align}
	q &= (8-2g) / (8+g)\ ,  \label{eq:struct2ei} \\
	\som &= 1\ .
\end{align}

It should be noted that although one can map structured jet light curves onto an energy injection model, there is no energy being added to the blast wave. Rather, the ``injection'' of energy is the additional flux from higher energy portions of the jet decelerating to the point the observer is in their beaming cone.

@@

This degeneracy is both useful and unfortunate.  The presence of an extended slowly decaying or rising afterglow is a largely model independent prediction of misaligned viewing and the presence of non-trivial jet structure.  Unfortunately, this same model independence makes it very difficult to distinguish between jet structure profiles, leaving both $E(\theta)$ and $\thobs$ uncertain.
